---
title: "WorryVisualSearch"
output: html_document
date: "2025-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Set enviorment and load packages 
```{r include=FALSE}

# clear the environment
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)


rawtrial<- read.csv("~/Desktop/Uni/project/data/trials_data_deidentified(in).csv")
rawsurvey <- read.csv("~/Desktop/Uni/project/data/survey_item_data.csv")
rawdemo <- read.csv("~/Desktop/Uni/project/data/demographics_data.csv")
# survey and trial
```

## 1) Select Task variables we want

Select the key variables: subject ID, practice vs main study, trial number, block, condition, participant response, correctness, and attention check indicator.

how the variables are recorded - make usre ot understand the datasets, how it is formatted 
when using visual studio code, check the ai generated code

each Ps will have thier worry group - look at that 

survey: basic demographics - sex age income education 
```{r select-variables}

# Select key trial variables
selected_trial <- rawtrial %>%
  select(
    subject_id = subID,                    # Subject ID
    uid = uid,                             # Unique ID
    session_id = sessionID,                # Practice vs main study identifier
    trial_number = trial_number,           # Trial number
    block = block_number,                  # Block number
    condition = condition,                 # Condition (target present/absent)
    task = task,                           # Task (practice vs actual)
    response = response,                   # Participant response
    correct = correct,                     # Correctness (1 = correct, 0 = incorrect)
    is_attention_check = is_attention_check, # Attention check indicator
    rt = rt,                               # Reaction time (useful for later analysis)
    worry_group = worryGroup               # Worry group classification
  )

# Display first few rows to verify selection
head(selected_trial)

# Summary of selected data
dim(selected_trial)
str(selected_trial)

# Select key survey variables
selected_survey <- rawsurvey %>%
  select(
    subject_id = participant_id,                    # Subject ID
    session_id = sessionID,                # Practice vs main study identifier
    survey_name = survey_name,             # Survey name/type
    question_name = question_name,         # Question identifier
    question_number = question_number,     # Question number in survey
    question_prompt = question_prompt,     # Full question text
    response = response,                   # Participant response
    response_label = response_label,       # Labeled response (e.g., "Strongly Agree")
    is_inattention_check = is_inattention_check, # Inattention check indicator
    worry_group = worryGroup,              # Worry group classification
    wdq_score = wdqScore,                  # Worry/Depression questionnaire score
    worry_cutoff = wdqCutoffUsed           # Whether worry cutoff was used
  )

# Display first few rows to verify selection
head(selected_survey)

# Summary of selected survey data
dim(selected_survey)
str(selected_survey)

```

## 2) Participant counts and trials per participant

Report how many unique participants are in the dataset and how many trials each participant completed.

make sure no missing data
general stats 

- condition count 
- exclude the practice tasks 
```{r participant-counts}
# Check for missing data in selected trial variables
missing_data <- selected_trial %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_count")

cat("Missing data summary: \n")
print(missing_data)

# Count unique participants
n_participants <- selected_trial %>%
  distinct(subject_id) %>%
  nrow()

cat("Total unique participants:", n_participants, "\n\n")


# Trials per participant
trials_per_participant <- selected_trial %>%
  group_by(subject_id) %>%
  summarise(
    n_trials = n(),
    n_blocks = n_distinct(block),
    n_conditions = n_distinct(condition),
    .groups = "drop"
  ) %>%
  arrange(desc(n_trials))

cat("Trials per participant summary:\n")
print(summary(trials_per_participant$n_trials))

cat("Trials per participant (first 10):\n")
print(head(trials_per_participant, 10))

# Overall statistics
cat("Overall dataset statistics:\n")
cat("Total trials:", nrow(selected_trial), "\n")
cat("Mean trials per participant:", round(mean(trials_per_participant$n_trials), 2), "\n")
cat("Min trials per participant:", min(trials_per_participant$n_trials), "\n")
cat("Max trials per participant:", max(trials_per_participant$n_trials), "\n")
cat("Median trials per participant:", median(trials_per_participant$n_trials), "\n")

```

## 3) Attention check trials: counts and performance

Describe how many attention-check trials are present and whether participants got them all correct.
accuract in passing attention - if did not pass then remove them 
```{r attention-checks}
# Filter attention check trials
attention_check_trials <- selected_trial %>%
  filter(is_attention_check == 1)

print("Attention Check Trials Summary:\n")
cat("Total attention check trials:", nrow(attention_check_trials), "\n")
cat("Unique participants with attention checks:", 
    n_distinct(attention_check_trials$subject_id), "\n\n")

# Performance on attention checks by participant
attention_check_performance <- attention_check_trials %>%
  group_by(subject_id) %>%
  summarise(
    n_attention_checks = n(),
    n_correct = sum(correct, na.rm = TRUE),
    accuracy = sum(correct, na.rm = TRUE) / n(),
    all_correct = all(correct == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(accuracy)

cat("Attention check accuracy by participant:\n")
print(attention_check_performance)
cat("\n")

# Summary statistics
cat("Attention check accuracy statistics:\n")
print(summary(attention_check_performance$accuracy))
cat("\n")

attention_check_performance <- attention_check_performance %>%
  left_join(
    selected_trial %>% select(subject_id, worry_group) %>% distinct(),
    by = "subject_id"
  )

# Plot attention check trial accuracy
ggplot(attention_check_performance, aes(x = worry_group, y = accuracy, fill = worry_group, colour = worry_group)) +
  geom_boxplot(alpha = 0.3, outlier.shape = NA) +
  geom_jitter(width = 0.1, alpha = 0.7, size = 2) +
  scale_fill_manual(values = c("high" = "#d9534f", "low" = "#5b9bd5")) +
  scale_colour_manual(values = c("high" = "#d9534f", "low" = "#5b9bd5")) +
  labs(
    title = "Accuracy on Attention Check Trials",
    x = "Worry Group",
    y = "Mean Accuracy",
    fill = "Worry Group",
    colour = "Worry Group"
  ) +
  theme_minimal()

# Identify participants who did NOT pass all attention checks
failed_attention <- attention_check_performance %>%
  filter(accuracy < 0.7)

cat("Number of participants who FAILED attention checks:", nrow(failed_attention), "\n")
if (nrow(failed_attention) > 0) {
  cat("Participants who failed attention checks:\n")
  print(failed_attention)
  cat("\n")
}

# Create cleaned dataset excluding participants who failed attention checks
participants_to_keep <- attention_check_performance %>%
  filter(accuracy > 0.7) %>%
  pull(subject_id)

selected_trial_clean <- selected_trial %>%
  filter(subject_id %in% participants_to_keep)

cat("Dataset after removing participants who failed attention checks:\n")
cat("Participants remaining:", n_distinct(selected_trial_clean$subject_id), "\n")
cat("Trials remaining:", nrow(selected_trial_clean), "\n")
cat("Trials removed:", nrow(selected_trial) - nrow(selected_trial_clean), "\n")


# Filter to only main task trials (exclude practice)
main_trials <- selected_trial_clean %>%
  filter(task == "visual_search", !is.na(condition)) %>%
  mutate(
    choice = case_when(
      condition == "condition_A" & response == "j" ~ "present",
      condition == "condition_A" & response == "f" ~ "absent",
      condition == "condition_B" & response == "f" ~ "present",
      condition == "condition_B" & response == "j" ~ "absent",
      TRUE ~ NA_character_
    ),
    acc = as.integer(correct == TRUE | correct == "True"),
    acc_factor = factor(ifelse(correct == TRUE | correct == "True", "Correct", "Incorrect"),
                        levels = c("Incorrect", "Correct")))
```


## 4) Per-participant accuracy and average reaction time

For each participant, create a table showing average accuracy and average reaction time.
- find average before cleaning

**Suggested accuracy cutoff:** A typical cutoff is 60% (0.60) to exclude low-performing participants, though 70% is also common depending on task difficulty.

```{r per-participant-summary}
# Calculate per-participant accuracy and RT (using cleaned data without failed attention checks)
# Exclude attention check trials from analysis

per_participant_summary <- main_trials %>%
  group_by(subject_id, worry_group, condition) %>%
  summarise(
    n_trials = n(),
    accuracy = mean(correct, na.rm = TRUE),
    mean_rt = mean(rt, na.rm = TRUE),
    sd_rt = sd(rt, na.rm = TRUE),
    min_rt = min(rt, na.rm = TRUE),
    max_rt = max(rt, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(accuracy))

cat("Per-participant accuracy and reaction time summary:\n")
print(per_participant_summary)
cat("\n")

# Summary statistics
cat("Accuracy statistics across participants:\n")
print(summary(per_participant_summary$accuracy))
cat("\n")

cat("Mean RT statistics across participants:\n")
print(summary(per_participant_summary$mean_rt))
cat("\n")

# Identify low performers (below cutoffs)
low_performers_60 <- per_participant_summary %>%
  filter(accuracy < 0.60)

low_performers_70 <- per_participant_summary %>%
  filter(accuracy < 0.70)

cat("Number of participants below 60% accuracy:", nrow(low_performers_60), "\n")
if (nrow(low_performers_60) > 0) {
  cat("Participants below 60% accuracy:\n")
  print(low_performers_60)
  cat("\n")
}

cat("Number of participants below 70% accuracy:", nrow(low_performers_70), "\n")
if (nrow(low_performers_70) > 0) {
  cat("Participants below 70% accuracy:\n")
  print(low_performers_70)
  cat("\n")
}

# Histogram of accuracy
hist(per_participant_summary$accuracy, 
     main = "Distribution of Per-Participant Accuracy",
     xlab = "Accuracy",
     ylab = "Frequency",
     breaks = 20,
     col = "steelblue",
     border = "black")
abline(v = 0.60, col = "red", lty = 2, lwd = 2, label = "60% cutoff")
abline(v = 0.70, col = "orange", lty = 2, lwd = 2, label = "70% cutoff")
legend("topleft", legend = c("60% cutoff", "70% cutoff"), 
       col = c("red", "orange"), lty = 2)

ggplot(per_participant_summary, aes(x = condition, y = accuracy, 
                                    group = subject_id, colour = worry_group)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.4) +
  scale_colour_manual(values = c("high" = "#d9534f", "low" = "#5b9bd5")) +
  labs(
    title = "Accuracy by Condition for Each Participant",
    x = "Condition",
    y = "Mean Accuracy",
    colour = "Worry Group"
  ) +
  theme_minimal()

# remove outliers
main_trials_clean <- main_trials %>%
  anti_join(low_performers_70, by = "subject_id")

# Histogram of mean reaction times -> convert ms to s
hist(per_participant_summary$mean_rt / 1000,      
     main = "Distribution of Per-Participant Mean Reaction Time",
     xlab = "Mean RT (s)",                        
     ylab = "Frequency",
     breaks = 20,
     col = "lightcoral",
     border = "black",
     xlim = c(0.8, 2.55))


# Calculate per-participant RT mean and SD for trial-level outlier removal
rt_per_participant <- main_trials %>%
  group_by(subject_id) %>%
  summarise(
    mean_rt = mean(rt, na.rm = TRUE),
    sd_rt   = sd(rt, na.rm = TRUE),
    .groups = "drop"
  )

# Join back to trial data and flag outlier *trials* (±2.5 or ±3 SD is common)
main_trials_clean <- main_trials %>%
  left_join(rt_per_participant, by = "subject_id") %>%
  filter(
    rt >= (mean_rt - 2.5 * sd_rt),
    rt <= (mean_rt + 2.5 * sd_rt)
  ) %>%
  select(-mean_rt, -sd_rt)  # drop helper columns

# Distribution of all trial RTs (converted ms → s)
hist(main_trials_clean$rt / 1000,                       
     main = "Distribution of Participant Reaction Time",
     xlab = "RT (s)",                              
     ylab = "Frequency",
     breaks = 20,
     col = "lightcoral",
     border = "black")




```
## 5) Consolidate All Excluded Participants

Combine all excluded participants from different criteria into one dataset for easy filtering of demographics using subject ID.

```{r excluded-participants-consolidation}
# Initialize a list to track excluded participants and reasons
excluded_list <- list()

# 1. Participants who failed attention checks (accuracy < 0.7)
failed_attention_ids <- attention_check_performance %>%
  filter(accuracy <= 0.7) %>%
  mutate(exclusion_reason = "Failed attention checks") %>%
  left_join(selected_trial %>% select(subject_id, uid) %>% distinct(), by = "subject_id") %>%
  select(subject_id, uid, exclusion_reason, accuracy_on_attention_checks = accuracy, worry_group)

if (nrow(failed_attention_ids) > 0) {
  excluded_list[[1]] <- failed_attention_ids
  cat("Excluded for failed attention checks:", nrow(failed_attention_ids), "\n")
}

# 2. Participants with low overall accuracy (< 70%)
low_performers_ids <- per_participant_summary %>%
  filter(accuracy < 0.70) %>%
  group_by(subject_id) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(exclusion_reason = "Low task accuracy (<70%)") %>%
  left_join(selected_trial %>% select(subject_id, uid) %>% distinct(), by = "subject_id") %>%
  select(subject_id, uid, exclusion_reason, mean_accuracy = accuracy, worry_group)

if (nrow(low_performers_ids) > 0) {
  excluded_list[[2]] <- low_performers_ids
  cat("Excluded for low accuracy:", nrow(low_performers_ids), "\n")
}

# Combine all exclusions into one dataset
excluded_participants <- bind_rows(excluded_list) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  arrange(subject_id)

# Create dataset of INCLUDED participants (for demographics filtering)
all_unique_participants <- selected_trial %>%
  distinct(subject_id, worry_group)

included_participants <- all_unique_participants %>%
  anti_join(excluded_participants, by = "subject_id")

cat("\n=== INCLUDED PARTICIPANTS ===\n")
cat("Total included participants:", n_distinct(included_participants$subject_id), "\n")
cat("Breakdown by worry group:\n")
print(table(included_participants$worry_group))

```

## 5.2) Attention check trials: survey & Demographics Filtering

```{r filter-demographics}
# Filter demographics data by:
# 1. Excluding participants from excluded_participants dataset
# 2. Including only participants with specified demographic criteria
demographics_filtered <- rawdemo %>%
  anti_join(excluded_participants, by = c("participant_id" = "uid")) %>%
  filter(
    dyslexia == "no",
    color_blindness == "no",
    fluent_english == "yes",
    vision %in% c("normal_vision", "corrected_to_normal_vision")
  )

# Summary of filtered demographics
cat("Demographics filtering summary:\n")
cat("Original participants in demographics:", n_distinct(rawdemo$participant_id), "\n")
cat("After excluding flagged participants:", 
    n_distinct(rawdemo %>% anti_join(excluded_participants, by = c("participant_id" = "uid")) %>% pull(participant_id)), "\n")
cat("Final after demographic criteria:", n_distinct(demographics_filtered$participant_id), "\n")

# Show what was excluded by demographic criteria
exclusions_summary <- rawdemo %>%
  anti_join(demographics_filtered, by = "participant_id") %>%
  summarise(
    total_excluded = n_distinct(participant_id),
    failed_dyslexia_filter = sum(dyslexia != "no", na.rm = TRUE),
    failed_colorblindness_filter = sum(color_blindness != "no", na.rm = TRUE),
    failed_english_filter = sum(fluent_english != "yes", na.rm = TRUE),
    failed_vision_filter = sum(!vision %in% c("normal_vision", "corrected_to_normal_vision"), na.rm = TRUE)
  )

# --- INCOME COMPLETENESS FILTER ---
# Identify participants with missing income before filtering
income_incomplete_ids <- demographics_filtered %>%
  filter(is.na(income)) %>%
  pull(participant_id)

cat("\nIncome completeness filter:\n")
cat("Participants missing income:", length(income_incomplete_ids), "\n")

# Exclude income-incomplete cases from demographics
demographics_filtered <- demographics_filtered %>%
  filter(!is.na(income))

cat("After excluding incomplete income:", n_distinct(demographics_filtered$participant_id), "\n")

# Propagate income exclusion to trial data
main_trials_clean <- main_trials_clean %>%
  filter(!uid %in% income_incomplete_ids)

# Filter out attention check items from survey data
selected_survey_clean <- selected_survey %>%
  filter(is_inattention_check == 0 | is.na(is_inattention_check))

cat("\nSurvey rows before filtering attention checks:", nrow(selected_survey), "\n")
cat("Survey rows after filtering attention checks:", nrow(selected_survey_clean), "\n")
cat("Attention check rows removed:", nrow(selected_survey) - nrow(selected_survey_clean), "\n")

# Identify subjects who FAILED survey attention checks
failed_survey_attn_ids <- selected_survey %>%
  filter(is_inattention_check == 1) %>%
  distinct(subject_id, worry_group) %>%
  mutate(
    exclusion_reason = "Failed survey attention check",
    subject_id = as.integer(subject_id)
  )

cat("\n=== SURVEY ATTENTION CHECK FAILURES ===\n")
cat("Subjects failed survey attention check:", nrow(failed_survey_attn_ids), "\n")
if (nrow(failed_survey_attn_ids) > 0) print(failed_survey_attn_ids)

# Add to the master excluded_participants list
excluded_participants <- bind_rows(
  excluded_participants,
  failed_survey_attn_ids %>%
    left_join(selected_trial %>% select(subject_id, uid) %>% distinct(), by = "subject_id")
) %>%
  distinct(subject_id, .keep_all = TRUE)

# Remove failed subjects from main_trials_clean and selected_survey_clean
main_trials_clean <- main_trials_clean %>%
  filter(!subject_id %in% failed_survey_attn_ids$subject_id)

selected_survey_clean <- selected_survey_clean %>%
  filter(!subject_id %in% failed_survey_attn_ids$subject_id)

cat("\n=== FINAL CLEAN DATASET AFTER SURVEY ATTENTION CHECK EXCLUSIONS ===\n")
cat("Participants remaining:", n_distinct(main_trials_clean$subject_id), "\n")
cat("Trials remaining:", nrow(main_trials_clean), "\n")
cat("Breakdown by worry group:\n")
print(table(distinct(main_trials_clean, subject_id, worry_group)$worry_group))

# 3. Participants with inconsistent WDQ worry group classification (pre vs post)
wdq_totals <- read.csv("data/wdq_totals_thesis.csv")

wdq_group_changers <- wdq_totals %>%
  filter(
    (worryGroup == "high" & wdq_post_total < 28) |
    (worryGroup == "low"  & wdq_post_total >= 28)
  )

cat("Participants excluded for WDQ group change:", nrow(wdq_group_changers), "\n")
print(wdq_group_changers)
```

## 5.3) Identify Participants Removed During Trial Cleaning

Compare original trial data with cleaned trial data to identify and document which participants were removed through the cleaning pipeline.

```{r removed-during-cleaning}
# All exclusions across demographics, survey, and trials
all_exclusions <- bind_rows(
  
  # 1. Failed trial attention checks
  failed_attention_ids %>%
    select(subject_id, uid, worry_group, exclusion_reason),
  
  # 2. Low task accuracy
  low_performers_ids %>%
    select(subject_id, uid, worry_group, exclusion_reason),
  
  # 3. Failed survey attention check
  failed_survey_attn_ids %>%
    left_join(selected_trial %>% select(subject_id, uid) %>% distinct(), by = "subject_id") %>%
    select(subject_id, uid, worry_group, exclusion_reason),
  
  # 4. Failed demographic criteria
  rawdemo %>%
    filter(
      dyslexia != "no" |
      color_blindness != "no" |
      fluent_english != "yes" |
      !vision %in% c("normal_vision", "corrected_to_normal_vision")
    ) %>%
    left_join(selected_trial %>% select(subject_id, uid, worry_group) %>% distinct(), 
              by = c("participant_id" = "uid")) %>%
    mutate(exclusion_reason = "Failed demographic criteria") %>%
    select(subject_id, uid = participant_id, worry_group, exclusion_reason),
  
  # 5. Incomplete income data
  rawdemo %>%
    filter(participant_id %in% income_incomplete_ids) %>%
    left_join(selected_trial %>% select(subject_id, uid, worry_group) %>% distinct(),
              by = c("participant_id" = "uid")) %>%
    mutate(exclusion_reason = "Incomplete income data") %>%
    select(subject_id, uid = participant_id, worry_group, exclusion_reason),
  
  # 6. WDQ worry group changed at retest
  wdq_group_changers %>%
    left_join(selected_trial %>% select(subject_id, uid, worry_group) %>% distinct(),
              by = c("participant_id" = "uid")) %>%
    mutate(exclusion_reason = "WDQ worry group changed at retest") %>%
    select(subject_id, uid = participant_id, worry_group, exclusion_reason)
  
) %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  arrange(subject_id)

cat("=== EXCLUSION LIST ===\n")
cat("Total unique excluded participants:", n_distinct(all_exclusions$subject_id), "\n\n")
cat("By reason:\n")
print(table(all_exclusions$exclusion_reason))
cat("\nBy worry group:\n")
print(table(all_exclusions$worry_group))
cat("\nFull list:\n")
print(all_exclusions)

main_trials_clean <- main_trials_clean %>%
  filter(!subject_id %in% all_exclusions$subject_id)

cat("Participants remaining:", n_distinct(main_trials_clean$subject_id), "\n")
cat("Trials remaining:", nrow(main_trials_clean), "\n")
```

## 6) Demographics Summary: Age, WDQ Score, and Education Level

Display distributions and summary statistics for key demographic variables in the filtered dataset.

```{r demographics-distributions}
demographics_filtered <- rawdemo %>%
  filter(participant_id %in% main_trials_clean$uid)

cat("=== DEMOGRAPHICS DISTRIBUTIONS FOR FILTERED SAMPLE ===\n\n")

# Summary statistics for age
cat("AGE Distribution:\n")
cat("N:", n_distinct(demographics_filtered$participant_id), "\n")
print(summary(demographics_filtered$age))
cat("SD:", sd(demographics_filtered$age, na.rm = TRUE), "\n\n")

# Summary statistics for WDQ score
cat("WDQ Score Distribution:\n")
print(summary(demographics_filtered$wdqScore))
cat("SD:", sd(demographics_filtered$wdqScore, na.rm = TRUE), "\n\n")

# Summary statistics for education level
cat("EDUCATION Level Distribution:\n")
print(table(demographics_filtered$education, useNA = "ifany"))
cat("\n")

# Histograms for age and WDQ score
par(mfrow = c(1, 2))

hist(demographics_filtered$age,
     main = "Distribution of Age",
     xlab = "Age (years)",
     ylab = "Frequency",
     col = "steelblue",
     border = "black",
     breaks = 10)

hist(demographics_filtered$wdqScore,
     main = "Distribution of WDQ Score",
     xlab = "WDQ Score",
     ylab = "Frequency",
     col = "lightcoral",
     border = "black",
     breaks = 10)

par(mfrow = c(1, 1))

# Bar plot for education level
education_counts <- demographics_filtered %>%
  group_by(education) %>%
  summarise(n = n_distinct(participant_id), .groups = "drop") %>%
  arrange(desc(n))

barplot(education_counts$n,
        names.arg = education_counts$education,
        main = "Distribution of Education Level",
        xlab = "Education Level",
        ylab = "Frequency",
        col = "lightgreen",
        border = "black")

# Display counts
print(education_counts)
```

## 7) WDQ Score by Worry Group

Compare WDQ scores between high and low worry groups with descriptive statistics and independent samples t-test.

```{r wdq-by-worry-group}
# Get worry group from the trial data to match with demographics
demographics_with_worry <- demographics_filtered %>%
  left_join(
    selected_trial %>% select(subject_id, uid, worry_group) %>% distinct(),
    by = c("participant_id" = "uid")
  )

# Summary statistics for WDQ scores by worry group
wdq_by_group <- demographics_with_worry %>%
  filter(!is.na(worry_group), !is.na(wdqScore)) %>%
  group_by(worry_group) %>%
  summarise(
    n = n(),
    mean_wdq = mean(wdqScore, na.rm = TRUE),
    sd_wdq = sd(wdqScore, na.rm = TRUE),
    .groups = "drop"
  )

cat("=== WDQ SCORE BY WORRY GROUP ===\n")
print(wdq_by_group)
cat("\n")

# Extract data for each group
high_worry_wdq <- demographics_with_worry %>%
  filter(worry_group == "high", !is.na(wdqScore)) %>%
  pull(wdqScore)

low_worry_wdq <- demographics_with_worry %>%
  filter(worry_group == "low", !is.na(wdqScore)) %>%
  pull(wdqScore)

# Perform independent samples t-test
t_test_result <- t.test(high_worry_wdq, low_worry_wdq)

cat("INDEPENDENT SAMPLES T-TEST:\n")
cat("High Worry Group: M = ", round(mean(high_worry_wdq, na.rm = TRUE), 2), 
    ", SD = ", round(sd(high_worry_wdq, na.rm = TRUE), 2), "\n", sep = "")
cat("Low Worry Group: M = ", round(mean(low_worry_wdq, na.rm = TRUE), 2), 
    ", SD = ", round(sd(low_worry_wdq, na.rm = TRUE), 2), "\n\n", sep = "")
cat("t(", round(t_test_result$parameter, 2), ") = ", round(t_test_result$statistic, 3), 
    ", p = ", round(t_test_result$p.value, 7), "\n", sep = "")
cat("\n")
```

## 8) Group-level accuracy by condition (present vs absent) - Table

Create a table showing mean accuracy for each condition.

 
```{r group-accuracy-table}

main_trials_clean %>%
  group_by(condition) %>%
  summarise(n = n()) %>%
  print()

mean_accuracy <- main_trials_clean %>%
  group_by(condition, worry_group) %>%
  summarise(
    mean_accuracy = mean(acc, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    condition = factor(
      condition,
      levels = c("target_absent", "target_present"),
      labels = c("Target Absent", "Target Present")
    )
  )

print(mean_accuracy)
```

## 8.2) Group-level accuracy by condition - Plot

Plot showing accuracy for target present vs absent conditions.
plot the error rates

```{r plot-accuracy-by-condition, fig.width=6, fig.height=4}
mean_accuracy <- main_trials_clean %>%
  group_by(condition, worry_group) %>%
  summarise(
    mean_accuracy = mean(acc, na.rm = TRUE),
    sd_accuracy = sd(acc, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    se_accuracy = sd_accuracy / sqrt(n),
    condition = factor(
      condition,
      levels = c("target_absent", "target_present"),
      labels = c("Target Absent", "Target Present")
    )
  )

participant_acc <- main_trials_clean %>%
  group_by(subject_id, condition, worry_group) %>%
  summarise(mean_accuracy = mean(acc, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    condition = factor(
      condition,
      levels = c("target_absent", "target_present"),
      labels = c("Target Absent", "Target Present")
    )
  )


ggplot(mean_accuracy, aes(x = condition, y = mean_accuracy, fill = worry_group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.6, alpha = 0.7) +
  geom_jitter(
    data = participant_acc,
    aes(y = mean_accuracy, colour = worry_group),
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.7),
    size = 2, alpha = 0.7
  ) +
  geom_errorbar(
    aes(ymin = mean_accuracy - se_accuracy, ymax = mean_accuracy + se_accuracy),
    position = position_dodge(width = 0.7),
    width = 0.4, linewidth = 0.8
  ) +
  coord_cartesian(ylim = c(0.6, 1.0)) +  
  scale_fill_manual(values = c("high" = "#E08080", "low" = "#80B0D0")) +
  scale_colour_manual(values = c("high" = "#C05050", "low" = "#5090B0")) +
  labs(
    title = "Accuracy by Condition and Worry Group",
    x = "Condition",
    y = "Mean Accuracy",
    fill = "Worry Group",
    colour = "Worry Group"
  ) +
  theme_minimal() +
  theme(legend.position = "right")

```

## 9) Group-level reaction time by condition - Table

Create a table summarising mean RT by condition. Consider excluding incorrect trials or extreme RTs.

```{r group-rt-table}
mean_rt <- main_trials_clean %>%
  group_by(condition, worry_group) %>%
  summarise(
    mean_rt = mean(rt, na.rm = TRUE)/1000,
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    condition = factor(
      condition,
      levels = c("target_absent", "target_present"),
      labels = c("Target Absent", "Target Present")
    )
  )

print(mean_rt)


```

## 9.2) Group-level reaction time by condition - Plot

Plot mean RT by condition with error bars.

```{r plot-rt-by-condition, fig.width=6, fig.height=4}
mean_rt <- main_trials_clean %>%
  group_by(condition,worry_group) %>%
  summarise(
     mean_rt = mean(rt, na.rm = TRUE) / 1000,
    sd_rt = sd(rt, na.rm = TRUE) / 1000,
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    se_rt = sd_rt / sqrt(n),
    condition = factor(
      condition,
      levels = c("target_absent", "target_present"),
      labels = c("Target Absent", "Target Present")
    )
  )

participant_rt <- main_trials_clean %>%
  group_by(subject_id, condition, worry_group) %>%
  summarise(mean_rt = mean(rt, na.rm = TRUE) / 1000, .groups = "drop") %>%
  mutate(
    condition = factor(
      condition,
      levels = c("target_absent", "target_present"),
      labels = c("Target Absent", "Target Present")
    )
  )

# Plot -- add error bars and subject means
ggplot(mean_rt, aes(x = condition, y = mean_rt, fill = worry_group)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(
    aes(ymin = mean_rt - se_rt, ymax = mean_rt + se_rt),
    width = 0.4,
    position = position_dodge(0.9)
  ) +
  geom_point(
    data = participant_rt,
    aes(x = condition, y = mean_rt, colour = worry_group),
    position = position_dodge(0.9),
    alpha = 0.5,
    size = 1.5
  ) +
  scale_fill_manual(values = c("high" = "#d9534f", "low" = "#5b9bd5")) +
  scale_colour_manual(values = c("high" = "#d9534f", "low" = "#5b9bd5")) +
  coord_cartesian(ylim = c(0, max(mean_rt$mean_rt) * 1.1)) +
  labs(
    title = "RT by Condition and Worry Group",
    x = "Condition",
    y = "Mean RT (s)",
    fill = "Worry Group"
  ) +
  guides(colour = "none") +
  theme_minimal()
```

## 10) Mixed Effects Logistic Regression Model for Accuracy

Fit mixed-effects logistic regression models with random intercept and slope for condition per participant.

```{r mixed-effects-setup, include=FALSE}
# Install packages if needed
if (!require("lme4")) install.packages("lme4")
if (!require("sjPlot")) install.packages("sjPlot")
if (!require("sjmisc")) install.packages("sjmisc")

library(lme4)
library(sjPlot)
library(sjmisc)
```

```{r prepare-data-for-lme}
# Prepare data for mixed effects model
# Join main_trials_clean with demographics_filtered to get demographic variables

lme_data <- main_trials_clean %>%
  left_join(
    demographics_filtered %>% select(participant_id, sex, age, education, income, wdqScore),
    by = c("uid" = "participant_id")
  ) %>%
  # Create scaled/centered variables
  mutate(
    # Scale trial number (z-score)
    trial_number_scaled = scale(trial_number)[,1],
    
    # Scale and center age
    age_z = scale(age)[,1],
    
    # Create previous trial condition (shift condition by 1 within each participant)
    prev_trial_condition = NA_character_,
    
    # Mean center condition (target_absent = -0.5, target_present = 0.5)
    condition_centered = ifelse(condition == "target_absent", -0.5, 0.5),
    
    # Mean center worry group (high = -0.5, low = 0.5)
    worry_group_centered = ifelse(worry_group == "high", -0.5, 0.5),
    
    # Ensure outcome variable is binary
    acc_binary = as.numeric(acc)
  ) %>%
  # Add previous trial condition
  group_by(subject_id) %>%
  mutate(prev_trial_condition = lag(condition)) %>%
  ungroup() %>%
  # Remove first trial for each participant (no previous trial)
  filter(!is.na(prev_trial_condition)) %>%
  # Convert factor variables as needed (keep education and income as numeric)
  mutate(
    condition = factor(condition, levels = c("target_absent", "target_present")),
    worry_group = factor(worry_group, levels = c("high", "low")),
    sex = factor(sex),
    prev_trial_condition = factor(prev_trial_condition, levels = c("target_absent", "target_present"))
  )

cat("Data prepared for mixed effects model:\n")
cat("N trials:", nrow(lme_data), "\n")
cat("N unique participants:", n_distinct(lme_data$subject_id), "\n")
cat("Check for missing data:\n")
print(colSums(is.na(lme_data[, c("acc_binary", "condition", "worry_group", "block", "trial_number_scaled", 
                                   "prev_trial_condition", "sex", "age_z", "education", "income")])))
```

```{r fit-lme-models}
# Filter to income-complete cases first
lme_data_income_complete <- lme_data %>% filter(!is.na(income))

# Base model on income-complete data
lme_accuracy_base <- glmer(
  acc_binary ~ condition * worry_group + block + trial_number_scaled +
    prev_trial_condition + sex + age_z + education +
    (1 + condition | subject_id),
  data = lme_data_income_complete,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e4))
)

print(summary(lme_accuracy_base))

# Full model on same data
lme_accuracy_full <- glmer(
  acc_binary ~ condition * worry_group + block + trial_number_scaled +
    prev_trial_condition + sex + age_z + education + income +
    (1 + condition | subject_id),
  data = lme_data_income_complete,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e4))
)

print(summary(lme_accuracy_full))

# Now anova() will work
cat("\n=== MODEL COMPARISON ===\n")
model_comparison <- anova(lme_accuracy_base, lme_accuracy_full)
print(model_comparison)
```

```{r lme-model-table}
# Display models in formatted table
tab_model(
  lme_accuracy_base,
  lme_accuracy_full,
  dv.labels = c("Accuracy (Base Model)", "Accuracy (Full Model with Income)"),
  show.stat = TRUE,
  show.aic = TRUE,
  show.loglik = TRUE,
  pred.labels = list(
    "Intercept",
    "Condition (Target Present)",
    "Worry Group (Low)",
    "Block",
    "Trial Number (Scaled)",
    "Prev Trial Condition (Target Present)",
    "Sex",
    "Age (Z-scored)",
    "Education",
    "Income",
    "Condition × Worry Group"
  ),
  title = "Mixed Effects Logistic Regression: Accuracy Prediction"
)
```

```{r lme-diagnostics}
# Extract random effects
cat("=== RANDOM EFFECTS ===\n")
cat("Base Model Random Effects:\n")
print(ranef(lme_accuracy_base))

# Plot random effects
par(mfrow = c(1, 2))
plot(ranef(lme_accuracy_base), main = "Base Model Random Effects")
plot(ranef(lme_accuracy_full), main = "Full Model Random Effects")
par(mfrow = c(1, 1))

# Fixed effects summary
cat("\n=== FIXED EFFECTS SUMMARY ===\n")
cat("Base Model Fixed Effects:\n")
print(fixef(lme_accuracy_base))

cat("\nFull Model Fixed Effects:\n")
print(fixef(lme_accuracy_full))
```

## 10.2) Mixed Effects Linear Regression Model for Reaction Time

Fit mixed-effects linear regression models for RT with random intercept and slope for condition per participant.

```{r prepare-data-for-lme-rt}
# Prepare data for RT mixed effects model
# Use the same lme_data but add RT variable

class(lme_data$rt)
str(lme_data$rt)
head(lme_data$rt)
lme_data %>%
  select(rt) %>%
  head()

lme_data_rt <- lme_data %>%
  left_join(
    main_trials_clean %>% select(subject_id, trial_number, block, rt),
    by = c("subject_id", "trial_number", "block")
  )

lme_rt <- lme_data %>%
  mutate(
    # Convert RT to seconds
    rt_seconds = rt / 1000,
    # Log-transform RT to normalize distribution
    log_rt = log(rt_seconds),
    # Scale log RT
    log_rt_scaled = scale(log_rt)[,1],
    # Keep education and income as numeric (not split into factor levels)
    education = as.numeric(education),
    income = as.numeric(income)
  ) %>%
  filter(!is.na(rt_seconds), !is.na(log_rt))

cat("Data prepared for RT mixed effects model:\n")
cat("N trials:", nrow(lme_rt), "\n")
cat("N unique participants:", n_distinct(lme_rt$subject_id), "\n")
cat("Mean RT (s):", round(mean(lme_rt$rt_seconds, na.rm = TRUE), 3), "\n")
cat("SD RT (s):", round(sd(lme_rt$rt_seconds, na.rm = TRUE), 3), "\n")
```

```{r fit-lme-models-rt}
# Base model: RT predicted by experimental factors (on log scale)

lme_rt_base <- lmer(
  log_rt ~ condition * worry_group + block + trial_number_scaled +
    prev_trial_condition + sex + age_z + education +
    (1 + condition | subject_id),
  data = lme_rt
)

cat("Base RT model fitted successfully\n")
print(summary(lme_rt_base))

# Full model: includes income
lme_rt_full <- lmer(
  log_rt ~ condition * worry_group + block + trial_number_scaled +
    prev_trial_condition + sex + age_z + education + income +
    (1 + condition | subject_id),
  data = lme_rt %>% filter(!is.na(income))
)

cat("\nFull RT model fitted successfully\n")
print(summary(lme_rt_full))

# Model comparison
# Create a dataset with no NAs on any variable used in either model
lme_rt_complete <- lme_rt %>%
  filter(
    !is.na(income),
    !is.na(sex),
    !is.na(age_z),
    !is.na(education),
    !is.na(log_rt_scaled)
  )

# Fit both models on the same dataset
lme_rt_base <- lmer(log_rt_scaled ~ condition * worry_group + block + 
                      trial_number_scaled + prev_trial_condition + 
                      sex + age_z + education + (1 + condition | subject_id),
                    data = lme_rt_complete)

lme_rt_full <- lmer(log_rt_scaled ~ condition * worry_group + block + 
                      trial_number_scaled + prev_trial_condition + 
                      sex + age_z + education + income + (1 + condition | subject_id),
                    data = lme_rt_complete)

cat("\n=== MODEL COMPARISON ===\n")
model_comparison_rt <- anova(lme_rt_base, lme_rt_full)
print(model_comparison_rt)
```

```{r lme-model-table-rt}
# Display RT models in formatted table
tab_model(
  lme_rt_base,
  lme_rt_full,
  dv.labels = c("RT - Base Model", "RT - Full Model with Income"),
  show.stat = TRUE,
  show.aic = TRUE,
  show.loglik = TRUE,
  pred.labels = list(
    "Intercept (log RT)",
    "Condition (Target Present)",
    "Worry Group (Low)",
    "Block",
    "Trial Number (Scaled)",
    "Prev Trial Condition (Target Present)",
    "Sex",
    "Age (Z-scored)",
    "Education",
    "Income",
    "Condition × Worry Group"
  ),
  title = "Mixed Effects Linear Regression: Reaction Time Prediction (Log Scale)"
)
```

```{r lme-diagnostics-rt}
# Extract random effects
cat("=== RANDOM EFFECTS FOR RT MODELS ===\n")
cat("Base Model Random Effects:\n")
print(ranef(lme_rt_base))

# Plot random effects
par(mfrow = c(1, 2))
plot(ranef(lme_rt_base), main = "Base Model Random Effects (RT)")
plot(ranef(lme_rt_full), main = "Full Model Random Effects (RT)")
par(mfrow = c(1, 1))

# Fixed effects summary
cat("\n=== FIXED EFFECTS SUMMARY FOR RT ===\n")
cat("Base Model Fixed Effects:\n")
print(fixef(lme_rt_base))

cat("\nFull Model Fixed Effects:\n")
print(fixef(lme_rt_full))

# Residual diagnostics
cat("\n=== RESIDUAL DIAGNOSTICS ===\n")
par(mfrow = c(2, 2))
plot(lme_rt_base, main = "Base Model Residuals")
qqnorm(residuals(lme_rt_base), main = "Q-Q Plot: Base Model")
qqline(residuals(lme_rt_base))
plot(lme_rt_full, main = "Full Model Residuals")
qqnorm(residuals(lme_rt_full), main = "Q-Q Plot: Full Model")
qqline(residuals(lme_rt_full))
par(mfrow = c(1, 1))

# Check for outliers
cat("\nResiduals Summary (Base Model):\n")
print(summary(residuals(lme_rt_base)))

cat("\nResiduals Summary (Full Model):\n")
print(summary(residuals(lme_rt_full)))
```
### 11) DDM m14 Analysis 
```{r m14-main-effects}
m14 <- read.csv("ddm/m14.csv", stringsAsFactors = FALSE)
 
# Standardise column names (in case they differ slightly)
names(m14) <- trimws(names(m14))
names(m14) <- gsub("%", "", names(m14))          # HDI 3% -> HDI 3
names(m14) <- gsub("\\s+", "_", names(m14))      # HDI 3 -> HDI_3
names(m14) <- gsub("\\.", "_", names(m14))
 
# Rename to standard internal names
m14 <- m14 %>%
  rename(
    parameter = matches("(?i)param"),
    mean      = matches("(?i)^mean$"),
    sd        = matches("(?i)^sd$"),
    hdi_lo    = matches("(?i)hdi_3"),
    hdi_hi    = matches("(?i)hdi_97"),
    ess       = matches("(?i)ess"),
    rhat      = matches("(?i)rhat|r_hat")
  ) %>%
  mutate(
    mean   = as.numeric(mean),
    hdi_lo = as.numeric(hdi_lo),
    hdi_hi = as.numeric(hdi_hi),
    sig    = !(hdi_lo < 0 & hdi_hi > 0)   # HDI excludes zero?
  )
 
# ---- 1. Extract individual parameters ----------------------
get_param <- function(name) {
  m14 %>% filter(parameter == name) %>% pull(mean)
}
 
a_int  <- get_param("a_Intercept")
a_tt   <- get_param("a_trialType")
a_wg   <- get_param("a_worryGroup")
a_int2 <- get_param("a_trialType:worryGroup")
v_int  <- get_param("v_Intercept")
v_tt   <- get_param("v_trialType")
```

```{r m14-cell-predictions}

# Sum-coded contrasts:
#   trialType  : -0.5 = Target Present, +0.5 = Target Absent
#   worryGroup : -0.5 = Low Worry,      +0.5 = High Worry
 
cells <- expand.grid(
  trialType  = c(-0.5, 0.5),
  worryGroup = c(-0.5, 0.5)
) %>%
  mutate(
    a_pred      = a_int + a_tt * trialType + a_wg * worryGroup +
                  a_int2 * trialType * worryGroup,
    v_pred      = v_int + v_tt * trialType,
    trial_label = ifelse(trialType  == -0.5, "Target Present", "Target Absent"),
    worry_label = ifelse(worryGroup == -0.5, "Low Worry",      "High Worry")
  )
 
pal <- c("High Worry" = "#E07B39", "Low Worry" = "#4A90D9")
```

```{r m14-boundary-seperation}
 
p_a <- ggplot(cells, aes(x = trial_label, y = a_pred,
                          colour = worry_label, group = worry_label)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 3.5) +
  scale_colour_manual(values = pal) +
  scale_x_discrete(limits = c("Target Present", "Target Absent")) +
  labs(
    title    = "Boundary Separation (a)",
    subtitle = "Higher = more evidence required before responding",
    x = NULL, y = "Predicted a (log scale)", colour = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title      = element_text(face = "bold"),
        legend.position = "bottom",
        panel.grid.minor = element_blank())

print(p_a)
```

``` {r m14-drift-rate}
p_v <- ggplot(cells, aes(x = trial_label, y = v_pred,
                          colour = worry_label, group = worry_label)) +
  geom_line(linewidth = 1.1, linetype = "dashed") +
  geom_point(size = 3.5) +
  scale_colour_manual(values = pal) +
  scale_x_discrete(limits = c("Target Present", "Target Absent")) +
  labs(
    title    = "Drift Rate (v)",
    subtitle = "No worry group effect — lines overlap by model design",
    x = NULL, y = "Predicted v", colour = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title      = element_text(face = "bold"),
        legend.position = "bottom",
        panel.grid.minor = element_blank())

print(p_v)
```

``` {r m14-forest-plot} 
library(patchwork)

forest_params <- m14 %>%
  filter(parameter %in% c("v_trialType", "a_trialType",
                           "a_worryGroup", "a_trialType:worryGroup")) %>%
  mutate(
    label = recode(parameter,
      "v_trialType"            = "v: Trial Type",
      "a_trialType"            = "a: Trial Type",
      "a_worryGroup"           = "a: Worry Group",
      "a_trialType:worryGroup" = "a: Trial Type \u00d7 Worry Group"
    ),
    hdi_label = paste0("[", round(hdi_lo, 3), ", ", round(hdi_hi, 3), "]")
  )
 
p_forest <- ggplot(forest_params,
                   aes(x = mean, y = reorder(label, mean), colour = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_errorbarh(aes(xmin = hdi_lo, xmax = hdi_hi),
                 height = 0.2, linewidth = 0.9) +
  geom_point(size = 4) +
  geom_text(aes(label = hdi_label),
            hjust = 0,        # left-align from point
            vjust = -0.8,     # lift above the point
            size = 3,
            colour = "grey30") +
  scale_colour_manual(
    values = c("TRUE" = "#2ECC71", "FALSE" = "#E74C3C"),
    labels = c("TRUE" = "HDI excludes 0", "FALSE" = "HDI includes 0")
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.2, 0.35))) +
  labs(
    title    = "Posterior Estimates: Group & Condition Effects",
    subtitle = "Mean \u00b1 94% HDI   |   ELPD LOO = \u22122929.5",
    x = "Posterior Mean (log-linear scale)", y = NULL, colour = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        legend.position  = "bottom",
        panel.grid.minor = element_blank())
print (p_forest)

combined <- (p_a | p_v) / p_forest +
  plot_annotation(
    title   = "DDM Full Interaction Model \u2014 Posterior Summary",
    caption = "Sum contrasts: Target Present = \u22120.5, Low Worry = \u22120.5.",
    theme   = theme(plot.title = element_text(size = 15, face = "bold"))
  )

ggsave("ddm_posterior_plots.png", combined,
       width = 14, height = 10, dpi = 180, bg = "white")

print(combined)

```
### 12) DDM m5 Analysis 
```{r m5-main-effects}
m5 <- read.csv("ddm/m5.csv", stringsAsFactors = FALSE)
 
# Standardise column names (in case they differ slightly)
names(m5) <- trimws(names(m5))
names(m5) <- gsub("%", "", names(m5))
names(m5) <- gsub("\\s+", "_", names(m5))
names(m5) <- gsub("\\.", "_", names(m5))
# Rename by position -- avoids regex failing on special characters like R̂
colnames(m5) <- c("parameter", "mean", "sd", "hdi_lo", "hdi_hi", "ess", "rhat")

m5 <- m5 %>%
  mutate(
    mean   = as.numeric(mean),
    hdi_lo = as.numeric(hdi_lo),
    hdi_hi = as.numeric(hdi_hi),
    sig    = !(hdi_lo < 0 & hdi_hi > 0)
  )
# ---- 1. Extract individual parameters ----------------------
get_param <- function(name) {
  m5 %>% filter(parameter == name) %>% pull(mean)
}
a_int <- get_param("a_Intercept")
a_tt  <- get_param("a_trialType")
v_int <- get_param("v_Intercept")
v_tt  <- get_param("v_trialType")


```

```{r m5-cell-predictions}
# M5 has trialType effects only -- no worryGroup parameter
# Sum-coded contrast: trialType: -0.5 = Target Present, +0.5 = Target Absent

cells <- data.frame(trialType = c(-0.5, 0.5)) %>%
  mutate(
    a_pred      = a_int + a_tt * trialType,
    v_pred      = v_int + v_tt * trialType,
    trial_label = ifelse(trialType == -0.5, "Target Present", "Target Absent")
  )

pal_trial <- c("Target Present" = "#4A90D9", "Target Absent" = "#E07B39")
```

```{r m5-boundary-seperation}
 
p_a5 <- ggplot(cells, aes(x = trial_label, y = a_pred,
                          colour = trial_label, group = 1)) +
  geom_line(linewidth = 1.1, colour = "grey50") +
  geom_point(size = 3.5, aes(colour = trial_label)) +
  scale_colour_manual(values = pal_trial) +
  scale_x_discrete(limits = c("Target Present", "Target Absent")) +
  labs(
    title    = "Boundary Separation (a)",
    subtitle = "Higher = more evidence required before responding",
    x = NULL, y = "Predicted a (log scale)", colour = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        legend.position  = "none",
        panel.grid.minor = element_blank())

print(p_a5)
```

``` {r m5-drift-rate}
p_v5 <- ggplot(cells, aes(x = trial_label, y = v_pred,
                          colour = trial_label, group = 1)) +
  geom_line(linewidth = 1.1, colour = "grey50") +
  geom_point(size = 3.5, aes(colour = trial_label)) +
  scale_colour_manual(values = pal_trial) +
  scale_x_discrete(limits = c("Target Present", "Target Absent")) +
  labs(
    title    = "Drift Rate (v)",
    subtitle = "Trial type effect only -- M5 contains no worry group parameter",
    x = NULL, y = "Predicted v", colour = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        legend.position  = "none",
        panel.grid.minor = element_blank())
print(p_v5)
```

``` {r m5-forest-plot} 
forest_params5 <- m5 %>%
  filter(parameter %in% c("v_trialType", "a_trialType")) %>%
  mutate(
    label = recode(parameter,
      "v_trialType" = "v: Trial Type",
      "a_trialType" = "a: Trial Type"
    ),
    hdi_label = paste0("[", round(hdi_lo, 3), ", ", round(hdi_hi, 3), "]")
  )

p_forest5 <- ggplot(forest_params,
                   aes(x = mean, y = reorder(label, mean), colour = sig)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_errorbarh(aes(xmin = hdi_lo, xmax = hdi_hi),
                 height = 0.2, linewidth = 0.9) +
  geom_point(size = 4) +
  geom_text(aes(label = hdi_label),
            hjust = 0, vjust = -0.8, size = 3, colour = "grey30") +
  scale_colour_manual(
    values = c("TRUE" = "#2ECC71", "FALSE" = "#E74C3C"),
    labels = c("TRUE" = "HDI excludes 0", "FALSE" = "HDI includes 0")
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.2, 0.35))) +
  labs(
    title    = "Posterior Estimates: Trial Type Effects (M5)",
    subtitle = "Mean \u00b1 94% HDI   |   Trial type only model",
    x = "Posterior Mean (log-linear scale)", y = NULL, colour = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        legend.position  = "bottom",
        panel.grid.minor = element_blank())

combined5 <- (p_a5 | p_v5) / p_forest5 +
  plot_annotation(
    title   = "DDM M5: Trial Type Effects -- Posterior Summary",
    caption = "Sum contrast: Target Present = -0.5, Target Absent = +0.5.",
    theme   = theme(plot.title = element_text(size = 15, face = "bold"))
  )

print(combined5)

ggsave("ddm_m5_posterior_plots.png", combined5,
       width = 14, height = 10, dpi = 180, bg = "white")

```

## 13) Factor Analysis for symptom dimensions
```{r rq2-score-questionnaires}
library(psych)

ddm_raw <- read.csv("ddm/subjectEffects_M5.csv")%>%
  rename(participant_id = subjectid)

#ddm_clean <- ddm_raw %>%
  #filter(participant_id %in% main_trials_clean$uid)

survey_score <- read.csv("data/survey_score_data.csv")
survey_analytic <- survey_score %>%
  filter(participant_id %in% ddm_raw$participant_id) %>%
  mutate(ocir_total = ocir_mean *18)

efa_vars <- survey_analytic %>%
  select(
    pswq_total,     # Penn State Worry Questionnaire
    gad7_total,     # Generalised Anxiety Disorder-7
    phq8_total,     # PHQ-8 (depression)
    ocir_mean      # OCI-R (obsessive-compulsive)
  )

cat("\nDescriptives:\n")
print(psych::describe(efa_vars))

cat("\nMissing values per variable:\n")
print(colSums(is.na(efa_vars)))

cat("\nMissing values per variable:\n")
print(colSums(is.na(efa_vars)))

# Listwise deletion
efa_complete <- efa_vars[complete.cases(efa_vars), ]

# Track which participant IDs survived listwise deletion (same order)
participant_ids_complete <- survey_analytic %>%
  filter(complete.cases(select(., gad7_total, phq8_total, ocir_total))) %>%
  pull(participant_id)

cat("\nN for EFA after listwise deletion:", nrow(efa_complete), "\n")
```

```{r rq2-efa-assumptions}
# Correlation matrix
cor_matrix <- cor(efa_complete, use = "complete.obs")
cat("Correlation matrix:\n", print(round(cor_matrix, 2)))

#Bartlett's test of spericity 
# Significant result = coreelation matrix is not an identity matrix (shared variance exists)
psych::cortest.bartlett(cor_matrix, n = nrow(efa_complete))

# Kaiser-Meyer-Olkin measure of sampling 
psych::KMO(efa_complete)
```

``` {r rq2-efa-parallel}
set.seed(41)
parallel_out <- psych::fa.parallel(
  efa_complete,
  fm     = "ml",
  fa     = "fa",
  n.iter = 1000,
  sim    = TRUE,
  main   = "Parallel Analysis Scree Plot"
)

cat("\nParallel analysis suggested factors:", parallel_out$nfact, "\n")
```

``` {r rq2-efa-fit}
efa_1f <- psych::fa(
  efa_complete,
  nfactors = 1,
  fm       = "ml",
  rotate   = "none"
)

cat("=== 1-Factor EFA Solution ===\n")
print(efa_1f$loadings, cutoff = 0.3)

cat("\nCommunalities:\n")
print(efa_1f$communality)

cat("\nVariance explained:\n")
print(efa_1f$Vaccounted)
```

```{r rq2-efa-scores}
# Extract factor scores (standardised: mean = 0, SD = 1)
factor_scores_df <- as.data.frame(efa_1f$scores)
colnames(factor_scores_df) <- "Factor1"
factor_scores_df$participant_id <- participant_ids_complete

cat("Factor score summary:\n")
print(psych::describe(factor_scores_df$Factor1))

# Plot factor loadings
loadings_df <- data.frame(
  Measure = c("PSWQ", "GAD-7","PHQ-8", "OCI-R"),
  Loading = as.numeric(efa_1f$loadings[,1])
)

ggplot(loadings_df, aes(x = Measure, y = Loading)) +
  geom_col(fill = "steelblue", width = 0.5) +
  geom_hline(yintercept = 0.3, linetype = "dashed", colour = "grey40") +
  ylim(0, 1) +
  labs(
    title = "EFA Factor Loadings (1-Factor Solution)",
    x     = "Measure",
    y     = "Loading on Factor 1"
  ) +
  theme_minimal()
```

``` {r rq2-join}
# reshape DDM output 
ddm_wide <- ddm_raw%>%
  select(participant_id, parameter, mean) %>%
  pivot_wider(names_from = parameter, values_from = mean) 

cat("DDM parameters:", print(names(ddm_wide)))
cat("N Particiapnts in DDM output:", nrow(ddm_wide))

# Join factor scores with DDM parameters
rq2_data <- factor_scores_df %>%
  inner_join (ddm_wide, by = "participant_id")

cat("\nN participants in final RQ2 dataset:", nrow(rq2_data), "\n")
cat("\nMissing values:\n", print(colSums(is.na(rq2_data))))

cat("\nDDM parameter descriptives:\n")
rq2_data %>%
  select(a, a_trialType, v, v_trialType, t, z) %>%
  psych::describe() %>%
  print()
```

``` {r rq2-regressions}
# Regress each DDM parameter onto the symptom factor score
# Primary parameter of interest is a_trialType(Boundary sep effect), followed by a (overall boundary sep)
reg_a         <- lm(a          ~ Factor1, data = rq2_data)
reg_a_trial   <- lm(a_trialType ~ Factor1, data = rq2_data)
reg_v         <- lm(v          ~ Factor1, data = rq2_data)
reg_v_trial   <- lm(v_trialType ~ Factor1, data = rq2_data)
reg_t         <- lm(t          ~ Factor1, data = rq2_data)

cat("=== Boundary Separation (a) ===\n");        print(summary(reg_a))
cat("=== Boundary Sep. x Condition ===\n");      print(summary(reg_a_trial))
cat("=== Drift Rate (v) ===\n");                 print(summary(reg_v))
cat("=== Drift Rate x Condition ===\n");         print(summary(reg_v_trial))
cat("=== Non-Decision Time (t) ===\n");          print(summary(reg_t))
```

```{r rq2-results-table}
# Formatted regression table
tab_model(
  reg_a, reg_a_trial, reg_v, reg_v_trial, reg_t,
  dv.labels = c(
    "Boundary Sep. (a)",
    "Boundary Sep. × Condition",
    "Drift Rate (v)",
    "Drift Rate × Condition",
    "Non-Decision Time (t)"
  ),
  pred.labels = c("Intercept", "Symptom Factor 1"),
  show.stat   = TRUE,
  show.fstat  = TRUE,
  title = "RQ2: Regression of DDM Parameters onto Transdiagnostic Symptom Factor"
)
```

```{r rq2-plot, fig.width = 9, fig.height = 5}
# Scatter plots: Factor 1 against each DDM parameter
rq2_long <- rq2_data %>%
  select(participant_id, Factor1, a, a_trialType, v, v_trialType, t) %>%
  pivot_longer(
    cols      = c(a, a_trialType, v, v_trialType, t),
    names_to  = "parameter",
    values_to = "estimate"
  ) %>%
  mutate(parameter = factor(parameter,
    levels = c("a", "a_trialType", "v", "v_trialType", "t"),
    labels = c("Boundary Sep. (a)", "Boundary Sep. × Condition",
               "Drift Rate (v)", "Drift Rate × Condition",
               "Non-Decision Time (t)")
  ))

ggplot(rq2_long, aes(x = Factor1, y = estimate)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = TRUE, colour = "steelblue", linewidth = 0.8) +
  facet_wrap(~ parameter, scales = "free_y", ncol = 3) +
  labs(
    title = "DDM Parameters Regressed onto Transdiagnostic Symptom Factor",
    x     = "Symptom Factor 1 Score",
    y     = "DDM Parameter Estimate"
  ) +
  theme_minimal()
```


